---
title: "Gender Equality DA"
author: "Sarah Shy"
date: "2/22/2020"
output: html_document
---

# Front matter
```{r}
# clear environment
rm(list = ls())

# load libraries
library(tidyverse)
```

```{r}
# load data
math.pisa <- read.csv("oecd_pisa_math.csv")
sci.pisa <- read.csv("oecd_pisa_science.csv")
grads <- read.csv("2017_grads_field.csv")
earn.deg <- read.csv("earnings_degree.csv")
earn.numeracy <- read.csv("earnings_numeracylevel.csv")
enroll <- read.csv("enrollment.csv")
```

For more data, see <https://stats.oecd.org/>

The data:
- Enrollment in STEM grad programs, broken down by gender, field, level of education
- PISA science scores  
- PISA math scores  
- Graduation broken down by degree, field, gender (2017)  
- Earnings broken down by degree, gender, age (values are percentages, 100 refers to the average earnings of all people who completed secondary school)  
- Earnings broken down by numeracy level, gender (all ages 25-64, data from 2012 and 2015)

Enrollment data comes from <https://stats.oecd.org/> table called "Enrolment by field"

Dataset documentation can be found [here (math)](https://data.oecd.org/pisa/mathematics-performance-pisa.htm)
and [here (science)](https://data.oecd.org/pisa/science-performance-pisa.htm#indicator-chart)

Share of graduates by field data comes from [here](https://stats.oecd.org/viewhtml.aspx?datasetcode=EAG_GRAD_ENTR_FIELD&lang=en#)

Earnings by numeracy level found [here](https://stats.oecd.org/viewhtml.aspx?datasetcode=EAG_EA_SKILLS&lang=en#)
